---
title: "Introduction to esctheme"
output: rmarkdown::html_vignette
description: Some example code showing how to use `theme61` with `ggplot2` graphs.
vignette: >
  %\VignetteIndexEntry{Introduction to esctheme}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r echo = FALSE, message = FALSE, warning = FALSE}
library(ggplot2)
library(data.table)
library(esctheme)

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.align = 'center', 
  out.width = '60%'
)

# No annoying messages
options(no_advisory = TRUE)

# Prep data: US personal expenditure aggregates in US$b
data <- USPersonalExpenditure
data <- as.data.table(data, keep.rownames = TRUE)
data <- melt(data, id.vars = "rn", variable.name = "year", value.name = "value")
setnames(data, "rn", "category")
data[, category := factor(category)]
data[, year := as.numeric(as.character(year))]

# 3 main categories
data_3 <- data[category %in% c("Food and Tobacco", "Household Operation", "Medical and Health")]

# Dual axis graph data
data_dual <- data[category %in% c("Food and Tobacco", "Personal Care")]
data_dual <- dcast(data_dual, year ~ category, value.var = "value")

# Column graph data
data_col <- data[year %in% c(1940, 1960) & category %in% c("Food and Tobacco", "Household Operation", "Medical and Health")]
data_col[, category := factor(category, 
                              c("Food and Tobacco", "Household Operation", "Medical and Health"),
                              c("Food", "Household", "Health"))]

# Change data
data_chg <- copy(data)[, change := value - shift(value), by = category]

# Point range data
pt_rg_data <- data.table(x = -2:1, 
                         ymin = c(-0.1, -0.2, 0, 0.2), 
                         y = c(0.2, 0.3, 0, 0.5), 
                         ymax = c(0.5, 0.8, 0, 0.8))


# Builds the colour palette diagram
inputs <- data.table::CJ(row = seq(1, 7), col = seq(0, 8))
inputs[, row := factor(row, labels = c("Blue", "Aqua", "Magenta", "Deepgreen", "Yellow", "Orange", "Red"))]

col_pal <- rbind(esc_blue, esc_aqua, esc_magenta, esc_deepgreen, esc_yellow, esc_orange, esc_red)

inputs[, fill_col := as.vector(t(col_pal))]

p <- ggplot(inputs, aes(x = col, y = row, fill = fill_col)) +
  geom_tile(show.legend = FALSE) +
  coord_fixed() +
  scale_fill_identity() +
  scale_x_continuous(n.breaks = 9) +
  labs(title = "The ESC Base Colour Palette") +
  esc_theme(rm_x_leg = TRUE, rm_y_leg = TRUE)

```


## Introduction

`esctheme` is the Essential Services Commission of Victoria graphing package for producing graphs that abide by our charts standards. It works by changing the appearance and functionality of charts produced using the `ggplot2` package.

You should read this article to help you understand how to create charts following our chrt standards.

This article provides some examples to help you figure out what elements you can change. The function documentation provides comprehensive information on the full customisation of the package.


## Correct workflow

When making graphs, you should follow the below workflow. 

Save your graph using `esc_save()`. Do **not** use the Export button in the Plots tab or `ggsave()`.

Start by creating a graph using the 'normal' stuff: `ggplot(data, aes(x, y)) + geom_blah()`. Then:

- Add plot labelling: `plot_label()`
- Use `theme_e61()` at the end of your plot code. 
- Save the graph: `esc_save()`.


## The colour palette

This is the Essential Services Commission colour palette. 

All the colours are package objects that can be selected by using the appropriate name. For example, `esc_magenta` selects the colour 'Magenta' with nine shades starting from the base shade in column '0'.

```{r theme61-colours, echo = FALSE}
p
```


## Graph examples

### Line graph {#line}

```{r}
g <- ggplot(data_3, aes(x = year, y = value, colour = category)) + 
  geom_line(linewidth = 1, show.legend = FALSE) +
  scale_colour_manual(values = esc_palette(3)) +
  plot_label(c("Food and\nTobacco", "Household\nOperation", "Medical and\nHealth"),
              c(1944, 1954, 1953),
              c(65, 55, 25),
             colour = esc_palette(3)
             ) +
  labs(
    title = "US Personal Expenditures*",
    subtitle = "1940-1960",
    caption = "* Data comes from the built-in US Personal Expenditure dataset."
  ) +
  ylab("US$ billions") +
  esc_theme(rm_x_leg = TRUE)

```

```{r, echo=FALSE}
#knitr::include_graphics("../man/figures/g-line.svg")
```

### Column graph {#column}

```{r}
g <- ggplot(data_3, aes(x = year, y = value, fill = category)) + 
  geom_col(position = "dodge", show.legend = FALSE, colour = "black") +
  scale_y_continuous(expand = expansion(mult = c(0,.2))) +
  scale_fill_manual(values = esc_palette(3)) +
  plot_label(c("Food and\nTobacco", "Household\nOperation", "Medical and\nHealth"),
              c(1936, 1943, 1950),
              c(85, 85, 85),
             colour = esc_palette(3)) +
  labs(
    title = "US Personal Expenditures*",
    subtitle = "1940-1960",
    caption = "* Data comes from the built-in USPersonalExpenditure dataset."
  ) +
  ylab("US$ billions") +
  esc_theme(rm_x_leg = TRUE)


```

```{r, echo=FALSE}
#knitr::include_graphics("../man/figures/g-col.svg")
```

### Scatter graph {#scatter}

```{r}

# Set up data
mtcars2 <- within(mtcars, {
  vs <- factor(vs, labels = c("V-shaped", "Straight"))
  am <- factor(am, labels = c("Automatic", "Manual"))
  cyl  <- factor(cyl)
  gear <- factor(gear)
})

# Graph
g <- ggplot(mtcars2, aes(x = wt, y = mpg, colour = gear)) +
  geom_point(size = 4, show.legend = FALSE) +
  scale_colour_manual(values = esc_palette(3)) +
  plot_label(
    label = c("3 cyl", "4 cyl", "5 cyl"),
    x = c(4, 2.7, 2.5),
    y = c(23, 35, 15),
    colour = esc_palette(3),
    size = 4.5
  ) +
  labs(
    title = "Car Fuel Efficiency and Weight",
    subtitle = "Data from 1973-74"
  ) +
  xlab("Weight (1000 lbs)") +
  ylab("Miles per gallon") +
  esc_theme()

```

```{r, echo=FALSE}
#knitr::include_graphics("../man/figures/g-scatter.svg")
```

### Faceted scatter graph {#facet}

```{r}
g <- ggplot(mtcars2, aes(x = wt, y = mpg)) +
  facet_wrap(~vs, scale = "free_x") +
  geom_point(size = 4, colour = esc_magenta[1], alpha = .7) +
  scale_y_continuous(limits = c(0, 40), breaks = seq(0,40, 12)) +
  labs(
    title = "Car Fuel Efficiency and Weight",
    subtitle = "Data from 1973-74"
  ) +
  xlab("Weight (1000 lbs)") +
  ylab("Miles per gallon") +
  esc_theme()

```

```{r, echo=FALSE, out.width='100%'}
#knitr::include_graphics("../man/figures/g-facet.svg")
```

### Density graph {#density}

```{r}
g <- ggplot(mpg, aes(x = cty, fill = factor(cyl))) + 
  geom_density(linetype = "blank", alpha = 0.7, show.legend = FALSE) +
  scale_fill_manual(values = esc_palette(4)) +
  scale_y_continuous(expand = 0) +
  
  labs(
    title = "Distribution of Car City Mileage",
    subtitle = "Miles per gallon, by number of cylinders"
  ) +
  plot_label(label = c("3 cyl", "5 cyl", "6 cyl", "8 cyl"),
              x = c(25, 24, 16, 12),
              y = c(0.1, 0.35, 0.25, 0.25),
             size = 4.5,
             colour = esc_palette(4)) +
  ylab("Density") +
  esc_theme(rm_x_leg = TRUE)

```

```{r, echo=FALSE}
#knitr::include_graphics("../man/figures/g-dens.svg")
```

### Horizontal bar graph {#flippedbar}



```{r}
g <- ggplot(data_col, aes(x = category, y = value, fill = factor(year))) +
  geom_col(position = "dodge") +
  coord_flip() +
  scale_fill_manual(values = esc_palette(2)) +
  plot_label(c("1940", "1960"),
              c(1.8, 2.2),
              c(15, 50),
             size = 4.5,
             colour = esc_palette(2)) +
  labs(
    title = "1940 to 1960 Change"
  ) +
  ylab("US$ billions") +
  esc_theme(rm_y_leg = TRUE) +
  format_flip()

```








